\documentclass{exercices}
\usepackage{ae, aeguill, graphicx}
\usepackage{fullpage}
\usepackage{xspace}
\usepackage{color}
\usepackage{amsmath, amssymb}
\usepackage{latexsym}
\usepackage{url}
\usepackage{tikz}
\usetikzlibrary{arrows,shapes,trees,fit,decorations.pathmorphing,backgrounds}

\usepackage{verbatim}

\begin{document}

\sujet{Java lab 2}

The answers to the exercices below are due before \emph{20 October 2008}. 
The procedure to send the exercices is:
\begin{enumerate}
  \item Create a directory named \verb!yourname-lab2!.
  \item Copy all the \verb!.java! source files to this directory.
  \item Issue following command in a shell : \verb!gtar czvf yourname-lab2.tar.gz yourname-lab2/!
        to create a tarball.
  \item Send this tarball to \verb!pablo@sifflez.org! with subject \verb!yourname-lab2!.
\end{enumerate}

\section{Goal}
This lab is divided in two sections:
\begin{itemize}
\item
In the first section you will write a java program that keeps a list of students and teachers, sorts
this persons' list by name and prints it.
\item
In the second section you will add to this program the possibility to keep track of courses, each course
will have a teacher and many students. You will add a class that tries to find a schedule so that
no teacher has to teach two courses the same day, and no student has to follow two courses the same day.
\end{itemize}

\section{People database}
\subsection{Classes hierarchy}
\begin{tikzpicture}
[classbox/.style={rectangle, draw=black!50, rectangle split,
  rectangle split parts=2, node distance=7cm, text width=6cm}]
  \node[classbox](Person)  {\verb!abstract class Person!\nodepart{second}
    \verb!Person(String firstName,!\newline
    \verb!       String lastName)!\newline
    \verb!String getFirstName()!\newline
    \verb!String getLastName()!\newline
    \verb!abstract void print()!
  };
\node[right of=Person,classbox](People)  {\verb!class People!\newline\verb!     extends ArrayList<Person>!\nodepart{second}
    \verb!void printList()!
  } edge[dashed] (Person);
\node[below left of=Person,classbox] (Student) {\verb!class Student!\nodepart{second}
  \verb!Student(String firstName,!\newline
  \verb!        String lastName, int year)! \newline
  } edge[<-] (Person);
\node[below right of=Person,classbox] (Teacher) {\verb!class Teacher!\nodepart{second}
  \verb!Teacher(String firstName,!\newline
  \verb!        String lastName)! \newline
  } edge[<-] (Person);
\end{tikzpicture}

\begin{exercice}[\textbf{class Person}]\\
Write an abstract class \verb!Person! that implements above methods.
To store the first and last name use two private instance \verb!String! variables.
\end{exercice}

\begin{exercice}[\textbf{classes Teacher and Student}]\\
Write \verb!Teacher! and \verb!Student! classes that extend \verb!Person!.
Students should have in addition to a first and last name, a field for their current year in school. 
\verb!Teacher! and \verb!Student! classes should implement method print which prints the person name, the person type 
(teacher or student) and in the case of students their current year.
\end{exercice}

\begin{exercice}[\textbf{class People}]\\
Write a class \verb!People! that extends \verb!java.util.ArrayList<Person>! class.
This class will keep the list of students and teachers.
It will provide a method \verb!printList!, which calls \verb!print()! for each \verb!Person! it contains.
Add a main method to this class that adds some teachers and students to \verb!People!, and call \verb!printList!.
\end{exercice}

\subsection{Sorting persons}
Now we want to sort the persons in \verb!People!, so that we can print them in alphabetical order.
To achieve this we will use method \verb!java.util.Collections.sort()! and interface \verb!java.util.Comparable!.
Read the documentation for interface \verb!Comparable! at
\url{http://java.sun.com/j2se/1.4.2/docs/api/java/lang/Comparable.html}
and for method sort at
\url{http://java.sun.com/j2se/1.4.2/docs/api/java/util/Collections.html#sort(java.util.List)}.

\begin{exercice}[\textbf{interface Comparable<Person>}]\\
We want to specify a total order for objects of type \verb!Person!. 
To do this we will modify class \verb!Person! so that it implements \verb!Comparable<Person>!.
We want to order persons using lexicographical order by their last name and if they have 
the same last name by their first name.

For example this would be a valid ordering:
\begin{verbatim}
  first   last

  Zoe     Amele
  Clement Dupont
  Cloe    Dupont
  Eric    Ziegler
\end{verbatim}
\end{exercice}
\begin{exercice}[\textbf{Sort People}]\\
Add a method sort to class \verb!People! that sorts its \verb!Persons!, using static method \verb!java.util.Collections.sort()!.
Test your implementation by calling \verb!printList!.
\end{exercice}
\section{Courses scheduler}
\begin{tikzpicture}
[classbox/.style={rectangle, draw=black!50, rectangle split,
  rectangle split parts=2, node distance=7cm, text width=7cm}]

  \node[classbox](Course)  {\verb!class Course!\newline
\verb!     extends ArrayList<Students>!
\nodepart{second}
    \verb!Course(String name,!\newline
    \verb!       Teacher teacher)!\newline
    \verb!int getDay()!\newline
    \verb!void setDay(int dayOfWeek)!\newline
    \verb!void unschedule()!\newline
    \verb!boolean scheduled()!\newline
    \verb!Teacher getTeacher()!\newline
    \verb!boolean compatible(Course other)!\newline
  };
\node[below of=Person,classbox] (Schedule) {\verb!class Schedule!\newline
\verb!     extends ArrayList<Course>!
\nodepart{second}
  \verb!static int daysInSchoolWeek = 5!\newline
  \verb!void print()!\newline
  \verb!void checkSchedule(Course course)!\newline
  \verb!     throws IncompatibleSchedule!\newline
  \verb!void updateSchedule()!\newline
  \verb!     throws IncompatibleSchedule,!\newline
  \verb!            FoundSchedule!
  } edge[dashed] (Course);
\end{tikzpicture}

\subsection{Courses}

\begin{exercice}[\textbf{Class Course}]\\
  Create a class course with the members listed above.
  Each \verb!Course! has a single teacher (provided in the constructor) and many students (added with method
  \verb!add! inherited from \verb!ArrayList!).
  \begin{itemize}
  \item
  Each course is either \emph{not scheduled} or \emph{scheduled for a specific day of the week, represented by an} \verb!int!.
  \item
  Method \verb!setDay!, allows to schedule a Course for a specific day.
  \item
  Method \verb!unschedule!, set the Course to \emph{not scheduled}.
  \item
  Method \verb!scheduled!, returns true if the course is scheduled.
  \item
  Method \verb!compatible!, check if current course is compatible with another course. ie, they have different teachers and they have no students in common.
  \end{itemize}
\end{exercice}

\subsection{Scheduler}
Finally we are going to write a class \verb!Schedule! that tries to find a compatible schedule for a list of courses.

\begin{exercice}[\textbf{Exceptions}]\\
Create two new exception classes called \verb!IncompatibleSchedule! and \verb!FoundSchedule!.
\end{exercice}


\begin{exercice}[\textbf{Schedule}]\\
Write a class \verb!Schedule! like above but only implement method \verb!print()!.
The method \verb!print! should print all the \verb!Courses! in \verb!Schedule!, and the day they are
scheduled or \emph{not scheduled} if they are not scheduled.
Add a list of courses to a \verb!Schedule! and call \verb!print()!, all the courses should be marked
as \emph{not scheduled}.
\end{exercice}

\begin{exercice}[\textbf{checkSchedule}]\\
Implement method \verb!checkSchedule!.
This method takes a \verb!Course!:
\begin{itemize}
\item if all the others courses scheduled the same day are compatible, it does nothing.
\item if one of them is not compatible, it throws exception \verb!IncompatibleSchedule!.
\end{itemize}
\end{exercice}

\begin{exercice}[\textbf{updateSchedule}]\\
Implement method \verb!updateSchedule!, using method \verb!checkSchedule!.
This method tries to find a schedule for all the \verb!Courses!, if this schedule is found it throws
\verb!ScheduleFound!, if no compatible schedule exists it throws \verb!IncompatibleSchedule!.

\begin{itemize}
\item To implement this method you will use a backtracking algorithm. 
\begin{enumerate}
  \item If all the courses are scheduled, throw \verb!ScheduleFound!.
  \item If there still are unscheduled courses, take one of them.
  \item For each of the available days:
  \begin{enumerate}
    \item Try to schedule if for day d.
    \item call \verb!updateSchedule()! recursively to try to schedule the other courses.
    \item if there is a schedule incompatibility, try another day.
  \end{enumerate}
  \item if there is no possible day on which to schedule the course, throw \verb!IncompatibleSchedule!. 
\end{enumerate}
\end{itemize}
\end{exercice}
\begin{exercice}[\textbf{Test}]\\
Add some students, teachers and courses and test the scheduling algorithm. 
To avoid adding a lot of courses, you can set \verb!daysInSchoolWeek! to a small value.
\end{exercice}
\end{document}
